---
title: "practice_exercise"
author: "Ekta Chaudhary"
date: "28/06/2020"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(sqldf)
library(readxl)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

#Reading the dataset

```{r}
practice_data = read_excel("./data/Practice_exercise.xlsx", sheet = "Data") %>%
  janitor::clean_names() %>%
  select(observation_number,quarter,employee_id, sex = sex_male_1, race, age, hospital_visit = hospital_visit_this_quarter_1_yes, salary, health_score) %>%
  mutate(
    age_cat = case_when(
      age < 30 ~ 1,
      age <= 45 ~ 2,
      age > 45 ~ 3
    )
  )
```

#Checking for missing data 

```{r}
sapply(practice_data, function(x) sum(is.na(x)))
```
```{r}
practice_data %>%
  select(everything()) %>%  
  summarise_all(funs(sum(is.na(.))))
```

#Finding the minimum and maximum values of each variable

```{r}
sapply(practice_data, function(x) min(x)) 
```
```{r}
sapply(practice_data, function(x) max(x)) 
```

#Checking the number of employees with health score outside the range of data

```{r}
practice_data %>%
  count(
   health_sc_6 = ifelse(health_score > 6, 1, 0)
  )
```
#Calculating the number of quarters for which the employees have missing data on sex

```{r}
practice_data %>%
  select(
    employee_id, sex
  ) %>%
  filter(
  is.na(sex)
  ) %>%
  group_by(
    employee_id
  ) %>%
  summarise(
    missing = sum(is.na(sex))
  )
```

#Calculating the number of quarters for which the employees have missing data on race

```{r}
practice_data %>%
  select(
    employee_id, race
  ) %>%
  filter(
     is.na(race)
  ) %>%
  group_by(
    employee_id
  ) %>%
  summarise(
    miss = sum(is.na(race))
  ) 
```

#Calculating the number of employees in each age group for each quarter

```{r}
emp_data = practice_data %>%
  mutate(
    quarter = factor(
    quarter),
    age_cat = factor(age_cat)
  )
```

```{r}
emp_data = emp_data %>%
  select(
    employee_id, quarter, age_cat
  ) %>%
  group_by(
    quarter, age_cat
  ) %>%
  tally()
```

```{r}
e <- ggplot(emp_data, aes(x = quarter, y = n, group = age_cat)) +
  geom_line(aes(color = age_cat)) +
  geom_point(aes(color = age_cat)) + labs(x = "Quarter", y = "Number of employees", title = "Number of employees in different age group for each quarter") + scale_color_discrete(labels = c("<30 years", "31-45 years", ">45 years"))
e
```

#Checking the trend in average salary over time 

```{r}
practice_data %>%
  select(
    salary, quarter
  ) %>%
  group_by(
    quarter
  ) %>%
  summarise(
    avg_salary = mean(salary)
  )
```

```{r}
salary_data = practice_data %>%
  mutate(
    age_cat = factor(age_cat),
quarter = factor(quarter))
```

#Checking the trend in average salary over time by age group

```{r}
salary_data = salary_data %>%
  select(
    salary, quarter, age_cat
  ) %>%
  group_by(
    quarter, age_cat
  ) %>%
  summarise(
    avg_salary = mean(salary)
  ) #%>%
  #pivot_wider(
    #names_from = age_cat, values_from = avg_salary
  #)
```

```{r}
p <- ggplot(salary_data, aes(x = quarter, y = avg_salary, group = age_cat)) +
  geom_line(aes(color = age_cat)) +
  geom_point(aes(color = age_cat)) + labs(x = "Quarter", y = "Average Salary", title = "Trend in average Salary over time by age group") + scale_color_discrete(labels = c("<30 years", "31-45 years", ">45 years"))
p
```

#Checking the trend in mean health score over time

```{r}
hc = practice_data %>%
  select(
    health_score, quarter
  ) %>%
  group_by(
    quarter
  ) %>%
  summarise(
    avg_score = mean(health_score)
  )
```

#Mean health score over time by age group

```{r}
practice_data %>%
  select(
    health_score, quarter, age_cat
  ) %>%
  group_by(
    quarter, age_cat
  ) %>%
  summarise(
    avg_score = mean(health_score)
  )
```

```{r}
health_sc = practice_data %>%
  mutate(
    age_cat = factor(age_cat),
quarter = factor(quarter))
```

```{r}
health_sc = health_sc %>%
  select(
    health_score, quarter, age_cat
  ) %>%
  group_by(
    quarter, age_cat
  ) %>%
  summarise(
    avg_score = mean(health_score)
  )
```

```{r}
p1 <- ggplot(health_sc, aes(x = quarter, y = avg_score, group = age_cat)) +
  geom_line(aes(color = age_cat)) +
  geom_point(aes(color = age_cat)) + labs(x = "Quarter", y = "Mean Health Score", title = "Trend in Mean Health Score over time by age group") + scale_color_discrete(labels = c("<30 years", "31-45 years", ">45 years"))
p1
```

#Mean health score over time by sex

```{r}
health_sex = practice_data %>%
  mutate(
    sex = factor(sex),
quarter = factor(quarter))
```

```{r}
health_sex = health_sex %>%
  drop_na() %>%
  select(
    health_score, quarter, sex
  ) %>%
  group_by(
    quarter, sex
  ) %>%
  summarise(
    avg_score = mean(health_score)
  )
```

```{r}
pq <- ggplot(health_sex, aes(x = quarter, y = avg_score, group = sex)) +
  geom_line(aes(color = sex)) +
  geom_point(aes(color = sex)) + labs(x = "Quarter", y = "Mean Health Score", title = "Trend in Mean Health Score over time by Sex") + scale_color_discrete(labels = c("Females", "Males"))
pq
```

#Mean health score over time by race

```{r}
health_race = practice_data %>%
  mutate(
    race = factor(race),
quarter = factor(quarter))
```

```{r}
health_race = health_race %>%
  drop_na() %>%
  select(
    health_score, quarter, race
  ) %>%
  group_by(
    quarter, race
  ) %>%
  summarise(
    avg_score = mean(health_score)
  )
```

```{r}
pe <- ggplot(health_race, aes(x = quarter, y = avg_score, group = race)) +
  geom_line(aes(color = race)) +
  geom_point(aes(color = race)) + labs(x = "Quarter", y = "Mean Health Score", title = "Trend in Mean Health Score over time by Race") 
pe
```

#Correcting the data quality issues

```{r}
new_data = practice_data %>%
  drop_na() %>%
  filter(
    health_score <= 6, 
    age >= 14, age <= 75
  )
```

#Checking the trend in mean health score over time in the corrected data

```{r}
new_data %>%
  select(
    health_score, quarter
  ) %>%
  group_by(
    quarter
  ) %>%
  summarise(
    avg_score = mean(health_score)
  ) %>%
  knitr::kable()
```

#Mean health score over time by age group
```{r}
new_data %>%
  select(
    health_score, quarter, age_cat
  ) %>%
  group_by(
    quarter, age_cat
  ) %>%
  summarise(
    avg_score = mean(health_score)
  ) %>%
knitr::kable()
```
```{r}
health_score = new_data %>%
  mutate(
    age_cat = factor(age_cat),
quarter = factor(quarter))
```

```{r}
health_score = health_score %>%
  select(
    health_score, quarter, age_cat
  ) %>%
  group_by(
    quarter, age_cat
  ) %>%
  summarise(
    avg_score = mean(health_score)
  )
```

```{r}
pn <- ggplot(health_score, aes(x = quarter, y = avg_score, group = age_cat)) +
  geom_line(aes(color = age_cat)) +
  geom_point(aes(color = age_cat)) + labs(x = "Quarter", y = "Mean Health Score", title = "Trend in Mean Health Score over time by age group") + scale_color_discrete(labels = c("<30 years", "31-45 years", ">45 years"))
pn
```

#Calculating the trend in hospital visits over time

```{r}
hosp = practice_data %>%
  select(
    employee_id, hospital_visit, quarter
  ) %>%
  group_by(
    quarter
  ) %>%
  summarise(
    percent = (sum(hospital_visit)/n())*100
  )
```

```{r}
s <- ggplot(hosp, aes(x = quarter, y = percent, color = percent)) + theme_classic() + geom_line() + geom_point()  +
   scale_x_discrete(name = "Quarter", limits = c("1","2","3","4","5","6","7","8","9","10","11","12")) + 
  scale_y_continuous(name = "Percentage of employees visiting hospitals",
                           breaks = seq(0, 25, 5),
                           limits = c(0, 25)) + labs(x = "Quarter", y = "Percentage of employees visiting hospitals" , title = "Trend in Hospital visits over time") 
s
```

#Calculating the distribution by gender over time

```{r}
w = practice_data %>%
  drop_na() %>%
  mutate(
    sex = factor(sex)
  ) %>%
  select(
     employee_id, quarter, sex
  ) %>%
  group_by(
    quarter, sex
  ) %>%
summarise(n = n()) %>%
mutate(freq = n / sum(n)*100) 
w
```

```{r}
t = w %>%
  ggplot(aes(x = quarter, y = freq, fill = sex)) + geom_bar(stat = "identity", legend = c("Female", "Males")) + scale_x_discrete(name = "Quarter", limits = c("1","2","3","4","5","6","7","8","9","10","11","12")) + 
  scale_y_continuous(name = "Percentage",
                           breaks = seq(0, 100, 10),
                           limits = c(0, 100)) + labs(x = "Quarter", y = "Percentage" , title = "Distribution by Gender over time") + scale_fill_discrete(name = "Sex", labels = c("Females", "Males")) 
t
```

#Calculating the distribution by race over time

```{r}
j = practice_data %>%
  drop_na() %>%
  mutate(
    race = factor(race)
  ) %>%
  select(
     quarter, race
  ) %>%
  group_by(
    quarter, race
  ) %>%
  summarise(n = n()) %>%
mutate(freq = n / sum(n)*100) 
j
```

```{r}
k = j %>%
  ggplot(aes(x = quarter, y = freq, fill = race)) + geom_bar(stat = "identity") +
  theme_classic() + scale_fill_manual(values = c("dodgerblue4","skyblue","pink")) + scale_x_discrete(name = "Quarter", limits = c("1","2","3","4","5","6","7","8","9","10","11","12")) + 
  scale_y_continuous(name = "Percentage",
                           breaks = seq(0, 100, 10),
                           limits = c(0, 100)) + labs(x = "Quarter", y = "Percentage" , title = "Distribution by Race over time") 
k
```

#Scatterplot between salary and health score

```{r}
practice_data %>%
  filter(
    health_score <= 6
  ) %>%
  ggplot(
    aes(x = health_score, y = salary)
  ) + geom_point(aes(color = "yellow")) + labs(
    x = "Heath Score", y = "Salary", title = "Association between Salary and Health Score"
  ) + scale_x_discrete(name = "Quarter", limits = c("1","2","3","4","5","6","7","8","9","10","11","12")) 
                           
```

#Box-plot showing the Mean health score for both genders

```{r}
practice_data %>%
  drop_na() %>%
  filter(
    health_score <= 6
  ) %>%
  mutate(
    sex = factor(sex)
  ) %>%
  ggplot(
    aes(
      x = sex, y = health_score
    )
  ) + geom_boxplot()
```

#Checking the association between sex and health score using two sample t-test

```{r}
t.test(health_score ~ sex, data = new_data)
```

#Checking the association between hospital visit and health score using two sample t-test

```{r}
t.test(health_score ~ hospital_visit, data = new_data)
```

#Scatterplot showing the association between age and health score

```{r}
new_data %>%
  filter(
    age <= 60
  ) %>%
ggplot(
  aes(
    x = age, y = health_score
  )
) + geom_point(
  aes(color = "pink")
)
```

#Checking the association between race and health score

```{r}
av <- aov(health_score ~ race, data = new_data)
summary(av)
```

#Creating 2*2 tables

```{r}
 practice_data %>%
  filter(
    age_cat != 3
  ) %>%
  mutate(
    sal_cat = case_when(
      salary < 40000 ~ "First",
      salary > 40000 ~ "Second"
    )
  ) %>%
  janitor::tabyl(age_cat, sal_cat)
```

```{r}
 practice_data %>%
  filter(
    age_cat != 3
  ) %>%
  mutate(
    sal_cat = case_when(
      salary < 40000 ~ "First",
      salary > 40000 ~ "Second"
    )
  ) %>%
  group_by(age_cat,sal_cat) %>%
  summarise(
    n = n()
  ) %>%
  pivot_wider(
    names_from = sal_cat, values_from = n
  )
```
#General summaries 

```{r}
practice_data %>%
  group_by(
    quarter
  ) %>%
  summarise(
    n = n(),
    mean_age = mean(age),
    median_age = median(age),
    mean_salary = mean(salary),
    median_salary = median(salary)
    )
```



