---
title: "Costello Medical exercise"
author: "Ekta Chaudhary"
date: "28/06/2020"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(readr)
library(plyr)
library(StatMeasures)
library(ggplot2)
set.seed(1)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

```{r}
path <- '/Users/ektachaudhary/Documents/practice_ex_2/Data/'
files = list.files(path = path, pattern = "*.csv", full.names = TRUE)
```

#Reading the datasets
```{r}
data_csv = ldply(files, read_csv) %>%
  janitor::clean_names()
data_csv 
```

#Created a new variable, BMI which is weight in kg/ (height in m)^2

```{r}
data_csv = data_csv %>%
  mutate(
    age_cat = case_when(
      age <= 50 ~ '<50',
      age <= 65 ~ '51-65',
      age > 65  ~ '65+'
    ),
    bmi = weight/(height/100)^2
    ) 
```

#Checking for missing values

```{r}
sapply(data_csv, function(x) sum(is.na(x))) %>%
  knitr::kable()
```


#Imputing the missing values for alcohol, height, bmi with the mean

```{r}
data_csv$alcohol[is.na(data_csv$alcohol)] <- mean(data_csv$alcohol, na.rm = TRUE)
data_csv$height[is.na(data_csv$height)] <- mean(data_csv$height, na.rm = TRUE)
data_csv$bmi[is.na(data_csv$bmi)] <- mean(data_csv$bmi, na.rm = TRUE)
```

#Exploring the minimimum and maximum value for each variable

```{r}
data_csv %>%
  summarise(
    Min_Age = min(age),
    Min_Weight = min(weight),
    Min_Alcohol_intake = min(alcohol),
    Min_Height = min(height),
    Min_BMI = min(bmi)
  ) %>%
knitr::kable()
```

```{r}
data_csv %>%
  summarise(
    Max_Age = max(age),
    Max_Weight = max(weight),
    Max_Alcohol_intake = max(alcohol),
    Max_Height = max(height),
    Max_BMI = max(bmi)
  ) %>%
knitr::kable()
```

#Checking the mean values for each variable

```{r}
data_csv %>%
  summarise(
    Avg_Age = mean(age),
    Avg_Weight = mean(weight),
    Avg_Alcohol_intake = mean(alcohol),
    Avg_Height = mean(height),
    Avg_BMI = mean(bmi)
  ) %>%
knitr::kable()
```

#Checking the age distribution:

```{r}
data_csv %>% 
  ggplot(aes(x = age)) + geom_histogram(color = "black", fill = "lightblue")
```

#Chceking the height distribution:

```{r}
data_csv %>% 
  ggplot(aes(x = height)) + geom_histogram(color = "black", fill = "pink")
```

#Chceking the weight distribution:

```{r}
data_csv %>% 
  ggplot(aes(x = weight)) + geom_histogram(color = "black", fill = "dark green")
```

#Checking the BMI distribution

```{r}
data_csv %>% 
  ggplot(aes(x = bmi)) + geom_histogram(color = "black", fill = "steel blue")
```

#Studying the comorbidities i.e., Diabetes, Hypertension, Diabetes&Hypertenion

```{r}
data_csv %>%
  drop_na() %>%
  ggplot(
    aes(
      comorbidities
    )
  ) + geom_bar(width = 0.4, fill = "steelblue", position = "dodge") + labs(x = "Comorbidities", y = "Number of patients" , title = "Number of patients with comorbidities") 
```

#Number of prior treatment for cirrhosis

```{r}
data_csv %>%
  drop_na() %>%
  ggplot(
    aes(
      number_of_prior_treatments_for_cirrhosis
    )
  ) + geom_bar(width = 0.4, fill = "darkgreen", position = "dodge") + labs(x = "Prior treatments of cirrhosis", y = "Number of patients" , title = "Number of patients with prior treatments of cirrhosis") 
```

# Associations:

#HCC and age

```{r}
age_dis = data_csv %>%
  select(
    age_cat, hcc_at_5_years
  ) %>%
  group_by(
    age_cat, hcc_at_5_years
  ) %>%
    dplyr::summarize(
      n = n()
      ) %>%
dplyr::mutate(freq = n / sum(n)*100) 
```

```{r}
age_dis
```


```{r}
age_dis %>%
  ggplot(aes(x = age_cat, y = freq, fill = hcc_at_5_years)) + geom_bar(stat = "identity", width = 0.5) + scale_x_discrete(name = "Age Category", limits = c("<50","51-65","65+")) + 
  scale_y_continuous(name = "Percentage",
                           breaks = seq(0, 100, 10),
                           limits = c(0, 100)) + labs(x = "Age Category", y = "Percentage" , title = "Association between HCC and Age") + scale_fill_discrete(name = "HCC at 5 years", labels = c("NO", "YES")) 
```

#HCC and BMI 

```{r}
data_csv %>%
  mutate(
    hcc_at_5_years = factor(hcc_at_5_years)
  ) %>%
  ggplot(
    aes(
      x = hcc_at_5_years, y = bmi, color = hcc_at_5_years
    )
  ) + geom_boxplot()  + labs(x = "HCC at 5 years", y = "Body Mass Index" , title = "Association between HCC and BMI") + scale_fill_discrete(name = "HCC at 5 years", labels = c("NO", "YES")) 
```

#HCC and alcohol

```{r}
hcc_alc = data_csv %>% 
  select(
    hcc_at_5_years, alcohol, age_cat
  ) %>%
  drop_na() %>%
  group_by(
    hcc_at_5_years, age_cat
  ) %>%
  dplyr::summarise(
    alc = mean(alcohol)
  ) %>%
  dplyr::mutate(
    alc = round(alc, digits = 2)
  )
```


```{r}
ggplot(hcc_alc, aes(x = factor(age_cat), y = alc)) +
  geom_bar(aes(fill = hcc_at_5_years), position = "dodge", stat = "identity", width = .5) +
      geom_text(aes(label = alc, group = hcc_at_5_years), position = position_dodge(width = 0.5), vjust = -0.5) + labs(x = "Age Category", y = "Average alcohol intake", title = "Association between HCC and alcohol intake") + scale_fill_discrete(name = "HCC at 5 years", labels = c("NO", "YES")) 
```

#Below 50, alcohol consumption doesn't effect the likelihood of developning HCC. As the age increases,higher alcohol consumption leads to an increased chances of developing HCC

#Hcc and comorbidities

```{r}
data_csv %>%
  select(
    hcc_at_5_years, comorbidities
  ) %>%
  drop_na() %>%
  group_by(
    comorbidities,hcc_at_5_years
  ) %>%
  dplyr::summarize(
      n = n()
      ) %>%
dplyr::mutate(freq = n / sum(n)*100) %>%
  mutate(
    freq = round(freq, digits = 2)
  ) %>%
  ggplot(aes(x = comorbidities, y = freq, fill = hcc_at_5_years)) + geom_bar(stat = "identity", width = 0.5) +  labs(x = "Comorbidities", y = "Percentage of patients", title = "Association between HCC and Comorbidities") + theme(axis.text.x = element_text(angle = 90, vjust = 0.9, hjust = 0.9)) + scale_fill_discrete(name = "HCC at 5 years", labels = c("NO", "YES")) 
```

#HCC and number of prior treatments for cirrhosis

```{r}
data_csv %>%
  select(
    hcc_at_5_years, number_of_prior_treatments_for_cirrhosis
  ) %>%
  drop_na() %>%
  group_by(
    hcc_at_5_years, number_of_prior_treatments_for_cirrhosis
  ) %>%
  dplyr::summarise(
    c = n()
  ) %>%
  ggplot(aes(x = number_of_prior_treatments_for_cirrhosis, y = c)) +
  geom_bar(aes(fill = hcc_at_5_years), position = "dodge", stat = "identity", width = .5) +
      geom_text(aes(label = c, group = hcc_at_5_years), position = position_dodge(width = 0.5), vjust = -0.5) +
      labs(x = "Number of prior treatments for Cirrhosis", y = "Number of patients", title = "Association between HCC and number of prior treatments for Cirrhosis") + scale_fill_discrete(name = "HCC at 5 years", labels = c("NO", "YES")) 
```

```{r}
data_csv$alcohol_dec <- decile(vector = data_csv$alcohol, decreasing = TRUE)
```

```{r}
alc_dec = data_csv %>% 
  select(
    alcohol,alcohol_dec
  ) %>%
  drop_na() %>%
  group_by(
    alcohol_dec
  ) %>%
  dplyr::summarise(
    alc = mean(alcohol)
  ) %>%
  dplyr::mutate(
    alc = round(alc, digits = 2)
  )
```

```{r}
alc_dec %>%
  drop_na() %>%
  ggplot(
    aes(x = factor(alcohol_dec), y = alc)
  ) + geom_bar(width = 0.4, fill = "steelblue", position = "dodge", stat = "identity") +
  geom_text(aes(label = alc), vjust = -0.3, size = 3.5) + labs(x = "Deciles", y = "Average alcohol consumption units" , title = "Average alcohol consumption by each Decile" )
```












