---
title: "Practice 2"
author: "Ekta Chaudhary"
date: "28/06/2020"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(sqldf)
library(readxl)
library(rvest)
library(plotly)
library(broom)
library(modelr)
library(knitr)
library(readr)
library(plyr)
set.seed(1)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

```{r}
path <- '/Users/ektachaudhary/Documents/practice_ex_2/Data/'
files = list.files(path = path, pattern = "*.csv", full.names = TRUE)
```

#Reading the datasets
```{r}
dat_csv = ldply(files, read_csv) %>%
  janitor::clean_names()
dat_csv 
```

#Checking for missing values

```{r}
sapply(dat_csv, function(x) sum(is.na(x)))
```

#Exploring the minimimum and maximum value for each variable

```{r}
data_new = dat_csv %>%
  drop_na()
sapply(data_new, function(x) min(x)) 
sapply(data_new, function(x) max(x))
```

#Checking the mean values for each variable

```{r}
dat_csv %>%
  drop_na() %>%
  summarise(
    Avg_age = mean(age),
    Avg_weight = mean(weight),
    Avg_Alcohol_intake = mean(alcohol),
    Avg_Height = mean(height)
  ) %>%
knitr::kable()
```

#Chceking the age distribution:

```{r}
dat_csv %>% 
  ggplot(aes(x = age)) + geom_histogram(color = "black", fill = "lightblue")
```

#Chceking the height distribution:

```{r}
dat_csv %>% 
  ggplot(aes(x = height)) + geom_histogram(color = "black", fill = "pink")
```

#Chceking the weight distribution:

```{r}
dat_csv %>% 
  ggplot(aes(x = weight)) + geom_histogram(color = "black", fill = "light yellow")
```
#Studying the comorbidities i.e., Diabetes, Hypertension, Diabetes&Hypertenion

```{r}
dat_csv %>%
  drop_na() %>%
  ggplot(
    aes(
      comorbidities
    )
  ) + geom_bar(width = 0.4, fill = "steelblue", position = "dodge") + labs(x = "Comorbidities", y = "Number of patients" , title = "Number of patients with comorbidities") 
```

#Number of prior treatment for cirrhosis

```{r}
dat_csv %>%
  drop_na() %>%
  ggplot(
    aes(
      number_of_prior_treatments_for_cirrhosis
    )
  ) + geom_bar(width = 0.4, fill = "darkgreen", position = "dodge") + labs(x = "Prior treatments of cirrhosis", y = "Number of patients" , title = "Number of patients with prior treatments of cirrhosis") 
```

# Associations:

#HCC and age

```{r}
new_data = dat_csv %>%
  mutate(
    age_cat = case_when(
      age < 50 ~ 1,
      age <= 75 ~ 2,
      age > 75  ~ 3
    )) 
```


```{r}
age_dis = new_data %>%
  select(
    age_cat, hcc_at_5_years
  ) %>%
  group_by(
    age_cat, hcc_at_5_years
  ) %>%
    dplyr::summarize(
      n = n()
      ) %>%
dplyr::mutate(freq = n / sum(n)*100) 
```

```{r}
age_dis %>%
  ggplot(aes(x = age_cat, y = freq, fill = hcc_at_5_years)) + geom_bar(stat = "identity", width = 0.5) + scale_x_discrete(name = "Age Category", limits = c("1","2","3")) + 
  scale_y_continuous(name = "Percentage",
                           breaks = seq(0, 100, 10),
                           limits = c(0, 100)) + labs(x = "Age Category", y = "Percentage" , title = "Association between HCC and Age") + scale_fill_discrete(name = "HCC at 5 years", labels = c("NO", "YES")) 
```

# HCC and weight

```{r}
new_data %>%
  mutate(
    hcc_at_5_years = factor(hcc_at_5_years)
  ) %>%
  ggplot(
    aes(
      x = hcc_at_5_years, y = weight, fill = hcc_at_5_years
    )
  ) + geom_boxplot() + scale_fill_brewer(palette = "Blues") + theme_classic()
```

# HCC and height

```{r}
new_data %>%
  mutate(
    hcc_at_5_years = factor(hcc_at_5_years)
  ) %>%
  ggplot(
    aes(
      x = hcc_at_5_years, y = height, fill = hcc_at_5_years
    )
  ) + geom_boxplot() + scale_fill_brewer(palette ="RdBu") + theme_classic()
```

#HCC and alcohol

```{r}
hcc_alc = new_data %>% 
  select(
    hcc_at_5_years, alcohol, age_cat
  ) %>%
  drop_na() %>%
  group_by(
    hcc_at_5_years, age_cat
  ) %>%
  dplyr::summarise(
    alc = mean(alcohol)
  ) %>%
  dplyr::mutate(
    alc = round(alc, digits = 2)
  )
```
```{r}
ggplot(hcc_alc, aes(x = factor(age_cat), y = alc)) +
  geom_bar(aes(fill = hcc_at_5_years), position = "dodge", stat = "identity", width = .5) +
      geom_text(aes(label = alc, group = hcc_at_5_years), position = position_dodge(width = 0.5), vjust = -0.5) + labs(x = "Age Category", y = "Average alcohol intake", title = "Association between HCC and alcohol intake") 
```

#Hcc and comorbidities

```{r}
dat_csv %>%
  select(
    hcc_at_5_years, comorbidities
  ) %>%
  drop_na() %>%
  group_by(
    hcc_at_5_years, comorbidities
  ) %>%
  dplyr::summarise(
    c = n()
  ) %>%
  ggplot(aes(x = comorbidities, y = c)) +
  geom_bar(aes(fill = hcc_at_5_years), position = "dodge", stat = "identity", width = .5) +
      geom_text(aes(label = c, group = hcc_at_5_years), position = position_dodge(width = 0.5), vjust = -0.5) +  labs(x = "Comorbidities", y = "Number of patients", title = "Association between HCC and Comorbidities") + theme(axis.text.x = element_text(angle = 90, vjust = 0.9, hjust = 0.9))
```

#HCC and number of prior treatments for cirrhosis

```{r}
dat_csv %>%
  select(
    hcc_at_5_years, number_of_prior_treatments_for_cirrhosis
  ) %>%
  drop_na() %>%
  group_by(
    hcc_at_5_years, number_of_prior_treatments_for_cirrhosis
  ) %>%
  dplyr::summarise(
    c = n()
  ) %>%
  ggplot(aes(x = number_of_prior_treatments_for_cirrhosis, y = c)) +
  geom_bar(aes(fill = hcc_at_5_years), position = "dodge", stat = "identity", width = .5) +
      geom_text(aes(label = c, group = hcc_at_5_years), position = position_dodge(width = 0.5), vjust = -0.5) +
      labs(x = "Number of prior treatments for Cirrhosis", y = "Number of patients", title = "Association between HCC and number of prior treatments for Cirrhosis") 
```


